### 2017-04-13 paper2.tex

Inter-site comparison is the crux of UHI analysis, thus it is imperative that urban T\textsubscript{surf} measurements are representative of coherent urban patches and free from confounding influences (i.e. atmospheric and emissivity effects). Meta-analysis of air temperature (T\textsubscript{air}) UHI literature shows that these goals are rarely satisfied \cite{Stewart2011}. Given the relative difficulty in retrieving accurate, representative urban T\textsubscript{surf}, similar conclusions are likely for sUHI analysis. In spite of this fact, and the short period over which large-scale, generalizable methods for measuring urban T\textsubscript{surf} have been available, study of sUHI via TIR remote sensing has expanded significantly in the last twenty years \cite{Peng2012,Voogt2003}. Using the method described in \ref{paper1} we derive an 8-month climatology of hemispherical radiometric urban T\textsubscript{surf} (T\textsubscript{hem, r}) from time continuous near-ground upwelling TIR as measured from an inverted pyrgeometer. The method was developed to address and overcome biases inherent in traditional methods for urban T\textsubscript{surf} retrieval by providing temporally continuous, geometrically representative\footnote{A hemispherical view is not perfectly geometrically representative of urban surface geometry. This is best illustrated by visualizing an urban area from the perspective of a downward facing ‘fish-eye’ camera. Geometric sampling biases are a result of lens distortions and the sensor cosine response. However, by sampling the surface 3-dimensionally, T\textsubscript{hem, r} is more representative of urban geometry than conventional 2-dimensional views of the surface. The geometric representivity of urban T\textsubscript{hem, r} is discussed in section \ref{Methodological limitations and considerations}.} urban T\textsubscript{surf} under all-sky conditions for sUHI analysis. These measurements are often made as a part of the net radiation determination for urban energy balance studies.

\subsection{Bias in thermal remote sensing}
Geometric biases in remote sensing of the urban surface are a result of its 3-dimensional, convoluted structure. Compared to flat terrain, complex urban surface geometry modifies receipt of incoming solar radiation and traps a portion of reflected solar and outgoing terrestrial radiation. The resulting microscale spatiotemporal contrasts in urban T\textsubscript{surf} create a directional dependence in observed urban T\textsubscript{surf} when measured from conventional narrow-field-of-view (FOV) TIR remote sensing platforms. Thus, remote sensed urban T\textsubscript{surf} varies based on sensor FOV, viewing angle and direction, and sun-surface geometry - this directional dependence of urban surface temperature is termed "effective thermal anisotropy" \cite{Voogt1998a}. Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface and fail to capture this effect - leading to spatiotemporally variant directional biases of up to 10 \si{\kelvin} in observed urban T\textsubscript{surf} \cite{Voogt1995}. In general, geometric undersampling by a remote sensor in the nadir manifests as an overestimation of daytime T\textsubscript{surf} and an underestimation of nighttime T\textsubscript{surf} \cite{Adderley2015}. However, the magnitude and diurnal and seasonal behaviors of this bias are dependent on sensor viewing geometry and unique site characteristics (canyon height-to-width ratio, canyon orientation and materials, vegetation coverage, etc.). Thus, parameterization schemes to account for urban effective anisotropy are difficult to generalize across urban sites, sensor types, and sensor-surface geometries.

In addition to undersampling the urban surface, most thermal remote sensing platforms yield an instantaneous ‘snap-shot’ and cannot characterize time-continuous T\textsubscript{surf} patterns without sacrificing ground resolution. Temporal discontinuities in thermal remote sensing result in myriad of potential sources of bias over a wide range of time scales. Aerial and satellite thermal remote sensing require clear sky conditions (clouds are opaque with respect to thermal infrared radiation). Hence, long term satellite characterizations of sUHI are biased towards conditions that maximize urban-rural contrasts in T\textsubscript{surf}. This results in an overestimation of “all-sky” sUHI. At diurnal scales, satellite overpass cycles rarely coincide with sUHI maximums and are not standard across cities or platforms. Thus, analysis of sUHI patterns and magnitudes across cities and instrument platforms is difficult. At smaller time scales still, time-continuous analysis of urban T\textsubscript{surf} shows significant microscale (second to minute) fluctuations in temperature \cite{Christen2012}. Most thermal remote sensors provide instantaneous T\textsubscript{surf} (rather than temporally averaged) and are potentially contaminated by high-frequency microscale fluctuations in urban T\textsubscript{surf}. This is particularly salient in urban environments, where a large variety of fabric materials can produce significant directional contrasts in thermal admittance - and thus spatial variations in the magnitude of microscale fluctuations depending on the facet material types viewed by the sensor. The effect of this phenomenon on thermal remote sensing has not been extensively studied, however, the magnitude of microscale fluctuations in T\textsubscript{surf} is significant relative to a typical sUHI signal and thus constitutes a potentially large source of bias.

@@

First, surface-to-sensor path length and view factor geometries are calculated with the surface-sensor-sun urban model (SUM) \cite{Soux2004}. Using a digital building model, SUM calculates path lengths from the surface to the sensor and angular view factors by projecting the pyrgeometer FOV onto a simplified (orthogonal) representation of the surface, determining which points in the DBM are 'seen' by the sensor, and calculating the distance from each 'seen' point to the sensor. Path lengths are binned at 5\si{\degree} intervals and averaged over the azimuth angle. Finally, view factors are calculated for each 5\si{\degree} angular as each bin's proportion of the total view factor and normalized to unity.

Second, version 4.1 of the MODTRAN radiative transfer code \cite{Berk1987} is used to model spectral at-sensor radiances for each angular path length. Runs are initialized using profiles of T\textsubscript{air} and water vapor content collected concurrently at each time step. Spectral radiances are modeled over a bandpass of 1 to 2500 \si{cm^{-1}} at an emissivity of 0.95 and convolved by an extended dome transmittance function to ensure that modeled radiances replicate the actual remote sensed signal. Consultation with radiometer manufacturers made clear the need to extend the modeled bandpass beyond a typical longwave bandpass (approximately 250 - 2500 \si{cm^{-1}}), as silicone domed pyrgeometers are transmissive of radiation in much shorter wavenumber (longer wavelengths) than 250 \si{cm^{-1}} (40 \si{\micro\meter}).

@@

T\textsubscript{hem, r} are retrieved from radiation and meteorological data collected from December 2001 through July 2002 as a part of the Basel Urban Boundary Layer Experiment (BUBBLE) campaign conducted in Basel, Switzerland \cite{Rotach2005}. Urban data were observed from a tower located near city center at the "Basel Sperrstrasse" site. Rural reference data were observed at the "Lange Erlen" site approximately 6 \si{\kilo \meter} to the north east in the outskirts of the city of Basel. Urban and rural site morphologies and observed variables are described in Table \ref{morphbspr}.

\begin{table}[H]
	\centering
	\caption{A description of morphological and measured variables for urban and rural sites. Modified from Rotach et al., 2005 to include only relevant parameters \cite{Rotach2005}.}
	\label{morphbspr}
	\begin{tabular*}{\textwidth}{p{3.75cm} p{2.25cm}p{3.5cm}p{2.75cm}p{2.75cm}}
		\toprule 
		Site & Location & Morphological & Meteorological & Radiation \\ 
		& Height & Characteristics\footnote{} & Variables\footnote{} & [No of levels]\footnote{} \\ 	\midrule

		Basel Sperrstrasse \newline \textit{Urban} \textit{street canyon} \newline LCZ: 2 & 47.57\si{\degree} N \newline 7.60\si{\degree} E \newline 255 \si{\meter} a.s.l. & $z_H$ = 14.6  \si{\meter} \newline $\sigma_H $ = 6.9 \si{\meter} \newline H/W = 0.54 \newline $\lambda_C $ = 0.37\newline $\lambda_S$ = 1.0 \newline $\alpha$ = 11.0\% & T\textsubscript{air} [7] \newline H [7] \newline WV [12] \newline WD [1] \newline P [1] & L\textsubscript{up} [3] \newline L\textsubscript{down} [5] \newline S\textsubscript{up} [2] \newline S\textsubscript{down} [3] \\ 
		& & & & \\
		Lange Erlen \newline \textit{Rural parkland} \newline LCZ: B & 47.59\si{\degree} N \newline 7.65\si{\degree} E \newline 240 \si{\meter} a.s.l. & $\alpha$ = 21.4\%  & T\textsubscript{air} [4] \newline H [4] \newline WV [3] \newline WD [1] &  L\textsubscript{up} [1] \newline L\textsubscript{down} [1] \newline S\textsubscript{up} [1] \newline S\textsubscript{down} [1]  \\ 
		\bottomrule
	\end{tabular*} 
		\raggedright
		\textsuperscript{2} Morphological parameters for the Sperrstrasse site were calculated for a 250 \si{\meter} circular area surrounding the study sites using the method described in \cite{Grimmond1999}. $z_H$: average building height, $\sigma_H $: standard deviation of building height, $\lambda_P $: plan aspect ratio, $\lambda_C $: complete aspect ratio, H/W: local canyon height to width ratio, $\alpha$: surface albedo. \\
		\textsuperscript{3} H: humidity, WV: wind velocity, WD: wind direction, P: pressure. \\
		\textsuperscript{4} L\textsubscript{up}: upwelling longwave radiation, L\textsubscript{down}: downwelling longwave radiation, S\textsubscript{up}: upwelling shortwave radiation, S\textsubscript{down}: downwelling shortwave radiation.
\end{table}

@@

Figure \ref{heatsuhi} shows seasonal variance in diurnal patterns of mean monthly hemispherical sUHI magnitudes at 30 minute intervals for the eight month study period. sUHI displays significant seasonal variance, particularly by day, and is stronly controlled by day length and solar angle. Nighttime sUHI does not vary significantly across seasons.

Figure \ref{heatcluhi} shows seasonal variance in diurnal patterns of mean monthly clUHI calculated at 30 minute intervals for the eight month study period. clUHI development is strongly controlled by the timing of sunset/sunrise cycles and displays the largest seasonal variance in the hours immediately after sunset. Daytime clUHI does not vary significantly across seasons.

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{heatsuhi}
	\caption{A heatmap of mean half hourly hemispherical sUHI for each month calculated at 30-minute intervals over the eight month study period.}
	\label{heatsuhi}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_kdown}
	\caption{sUHI magnitude versus incoming solar radiation binned for clear sky (when the sum of daytime K\textsubscript{down} \textgreater 10000 \si{\watt \per \square \meter})and cloudy sky days (when the sum of daytime K\textsubscript{down} \textless 10000 \si{\watt \per \square \meter})}
	\label{meteo_kdown}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_hum}
	\caption{sUHI magnitude versus atmospheric water vapor content measured at 2 \si{\meter}. Similar patterns are observed when rural water vapor content is substituted.}
	\label{meteo_hum}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_wv}
	\caption{Bees nest plot}
	\label{meteo_wv}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_atuhi}
	\caption{Bees nest plot}
	\label{meteo_atuhi}
\end{figure}

@@

sUHI in Basel from complete, nadir, and hemispherical representations of T\textsubscript{surf} displays two diurnal peaks, the first in the late afternoon and a second just after sunset. Clear sky conditions amplify this phenomena. A late afternoon peak in sUHI is likely the result of the urban-rural contrasts in thermal admittance (and by relation, urban-rural soil moisture contrasts). The thermal admittance of many urban surfaces, particularly rooftops, tends to be lower than rural and vegetated environments \cite{Spronken-Smith1998} – particularly in temperate climates where rural soil moisture is largely maintained through summer. Lower urban thermal admittance manifests in a larger diurnal thermal amplitude and thus, greater daytime surface heating and nighttime surface cooling. Nadir and hemispherical views are biased towards low thermal admittance rooftop facets, and thus display large daytime peaks in sUHI.

A second peak in sUHI after sunset is the result of differences in relative cooling rates of urban and rural surfaces. In this case, nighttime sUHI is likely the result of canyon radiation trapping from obstructed canyon sky-view-factors. The second peak sUHI is largest on following clear sky conditions when the potential for significant differentials in urban-rural nocturnal cooling rates is maximized. A second peak is observed across all representations of urban T\textsubscript{surf} but is most apparent in complete and nadir representations of sUHI, which, for the Sperrstrasse site, sample T\textsubscript{road} to a greater extent than the hemispherical view. A weak sUHI is maintained through the nighttime as urban and rural cooling rates reach parity in the hours after sunset \cite{Oke2017}. Minimum sUHI is observed just after sunrise, where sUHI magnitudes are occasionally negative. At low solar angles, shading by 3-dimensional urban geometry suppresses heating rates for a large portion of urban surface area, reducing overall urban heating rates in early morning. As solar angles increase through the morning and more facets experience direct isolation, sUHI develops, with nadir and hemispherical representations displaying more rapid intensification than complete sUHI from a bias towards low thermal admittance and directly sunlit rooftop surfaces.

@@

Hemispherical sUHI displays significant seasonality, both in terms of its diurnal pattern and overall magnitude. Seasonal variance in sUHI is greatest by day, where large seasonal contrasts in cloud cover frequency and solar angle exert strong control on daytime heating rates and sUHI development. Peak sUHI occurs near the summer solstice, when days are longest and isolation on clear sky days is most intense. In winter, daytime sUHI is suppressed from a lower solar input and frequent cloud cover. This shifts and weakens peak sUHI in winter to shortly after sunset. Nighttime sUHI is relatively consistent across seasons, largely a product of contrasts in urban-rural cooling rates just after sunset resulting from radiation trapping by canyon geometry. Similar patterns are observed in seasonal variance of canopy layer UHI, albeit concentrated at different times of day.

In contrast to sUHI, clUHI is greatest just after sunset and is strongly controlled by sunset/sunrise timing. Peak clUHI occurs just after sunset across seasons, differing only in terms of magnitude, with a larger peak clUHI in summer. However, in this study, it should be note that summertime clUHI magnitudes in the hours just after sunset are slightly higher than expected \cite{Runnalls2000} – possibly the result of cold air draining from hilly terrain west of the rural site. Seasonality in clUHI results from seasonal contrasts in cloud cover frequency. Increased wintertime cloud cover reduces radiative losses by the canyon air volume after sunset, suppressing mean wintertime clUHI magnitudes. Daytime clUHI is consistently small across all seasons. Thus, although sUHI and clUHI are controlled by similar climatic controls and have a similar micro meteorological genesis, they do not display the same seasonal patterns. Thus, \textit{a surface urban heat island is not simply a stronger, more pronounced, canopy layer urban heat island.}

@@

sUHI is highly dependent on meteorological conditions, which account for much of the variation in daytime sUHI magnitudes. Conditions that foster the largest microscale contrasts in urban T\textsubscript{surf} also entail the largest sUHI magnitudes. Contra, conditions that suppress microscale contrasts in T\textsubscript{surf} show similarly reduced sUHI magnitudes. As such, incoming shortwave radiation and wind velocity display positive and negative relationships with sUHI magnitudes respectively. Under clear sky conditions, strong solar heating of the surfaec combined with differences in urban and rural thermal properties fosters the development of large daytime urban-rural thermal contrasts. Thus, sUHI\textsubscript{max} coincides with insolation\textsubscript{max}. Under cloudy conditions, decreased direct solar heating of the surface suppresses the development of urban-rural thermal contrasts, resulting in a reduced sUHI. $\Delta$T\textsubscript{hem, r - air} has a similar positive effect on sUHI, owing to its dependence on solar insolation.

As was found with correction magnitudes in Chapter \ref{paper1}, sUHI magnitudes are not strongly correlated with water vapor content as, in the absence of fog, humidity near the surface does not strongly influence surface heating or cooling rates in and of itself. Atmospheric transmittance of longwave radiation decreases sharply with small increases in water vapor when the atmosphere is relatively dry - shown in Figure \ref{humtest} - with decreasing variation in atmospheric transmittance as water vapor content increases from zero. Therefore, typical variations in urban and rural humidities have little effect on the fraction of outgoing available for heating of the canopy layer air volume through absorption of outgoing longwave radiation and for heating of the surface by atmospheric emission towards the surface. As a result, water vapor content has little effect on sUHI. Wind velocity displays a strong negative relationship with sUHI. sUHI is greatest under calm winds when near-surface mixing is minimal, fostering strong microscale and urban-rural contrasts in T\textsubscript{surf}. As wind velocity increases, sUHI is suppressed from increased turbulent heat losses.

\begin{table}[H]
	\centering
	\caption{A summary of the relationship between sUHI, meteorological variables, and clUHI.}
	\label{meteo_cont}
	\begin{tabular}{lc}
		\toprule 
		An increase in $x$, has a/an... & $y$ effect on sUHI magnitude. \\
		\midrule
		$\Delta$T\textsubscript{hem, r - air} &\textit{increasing} \\
		&\\
		T\textsubscript{hem, r} & \textit{increasing} \\
		&\\
		T\textsubscript{air} & \textit{increasing} \\
		&\\
		clUHI &\textit{increasing} \\
		&\\
		Incoming solar radiation & \textit{increasing}  \\
		& \\
		& \\
		Water vapor content & \textit{negligible} \\
	 	&\\
		Wind velocity &\textit{decreasing} \\
		\bottomrule
	\end{tabular} 
\end{table}

clUHI and sUHI develop at different times of day and, thus, have a complex diurnal relationship. By day, variation in clUHI is minimal and shows a weak positive correlation with sUHI. By night, clUHI shows greater variation and a stronger positive relationship with sUHI. Increasing sUHI with clUHI is largest on clear sky nights when urban-rural contrasts in air and surface cooling rates are strongest. This results in similar sUHI and clUHI development in the hours after sunset under clear sky conditions.

>>>>

Inter-site comparison is the crux of UHI analysis, thus it is imperative that urban T\textsubscript{surf} measurements are representative of coherent urban patches and free from confounding influences (i.e. atmospheric and emissivity effects). Meta-analysis of air temperature (T\textsubscript{air}) UHI literature shows that these goals are rarely satisfied \citep{Stewart2011}. Given the relative difficulty in retrieving accurate, representative urban T\textsubscript{surf}, similar conclusions are likely for sUHI analysis. In spite of this fact, and the short period over which large-scale, generalizable methods for measuring urban T\textsubscript{surf} have been available, study of sUHI via TIR remote sensing has expanded significantly in the last twenty years \citep{Peng2012,Voogt2003}. Using the method described in Chapter \ref{paper1} we derive an 8-month climatology of hemispherical radiometric urban T\textsubscript{surf} (T\textsubscript{hem, r}) from time continuous near-ground upwelling TIR as measured from an inverted pyrgeometer. The method was developed to address and overcome biases inherent in traditional methods for urban T\textsubscript{surf} retrieval by providing temporally continuous, geometrically representative\footnote{A hemispherical view is not perfectly geometrically representative of urban surface geometry. This is best illustrated by visualizing an urban area from the perspective of a downward facing ‘fish-eye’ camera. Geometric sampling biases are a result of lens distortions and the sensor cosine response. However, by sampling the surface 3-dimensionally, T\textsubscript{hem, r} is more representative of urban geometry than conventional 2-dimensional views of the surface. The geometric representivity of urban T\textsubscript{hem, r} is discussed in section \ref{Methodological limitations and considerations}.} urban T\textsubscript{surf} under all-sky conditions for sUHI analysis. These measurements are often made as a part of the net radiation determination for urban energy balance studies.

\subsection{Bias in thermal remote sensing}
Geometric biases in remote sensing of the urban surface are a result of its 3-dimensional, convoluted structure. Compared to flat terrain, complex urban surface geometry modifies receipt of incoming solar radiation and traps a portion of reflected solar and outgoing terrestrial radiation. The resulting microscale spatiotemporal contrasts in urban T\textsubscript{surf} create a directional dependence in observed urban T\textsubscript{surf} when measured from conventional narrow-field-of-view (FOV) TIR remote sensing platforms. Thus, remote sensed urban T\textsubscript{surf} varies based on sensor FOV, viewing angle and direction, and sun-surface geometry - this directional dependence of urban surface temperature is termed "effective thermal anisotropy" \citep{Voogt1998a}. Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface and fail to capture this effect - leading to spatiotemporally variant directional biases of up to 10 \si{\kelvin} in observed urban T\textsubscript{surf} \citep{Voogt1995}. In general, geometric undersampling by a remote sensor in the nadir manifests as an overestimation of daytime T\textsubscript{surf} and an underestimation of nighttime T\textsubscript{surf} \citep{Adderley2015}. However, the magnitude and diurnal and seasonal behaviors of this bias are dependent on sensor viewing geometry and unique site characteristics (canyon height-to-width ratio, canyon orientation and materials, vegetation coverage, etc.). Thus, parameterization schemes to account for urban effective anisotropy are difficult to generalize across urban sites, sensor types, and sensor-surface geometries.

In addition to undersampling the urban surface, most thermal remote sensing platforms yield an instantaneous ‘snap-shot’ and cannot characterize time-continuous T\textsubscript{surf} patterns without sacrificing ground resolution. Temporal discontinuities in thermal remote sensing result in myriad of potential sources of bias over a wide range of time scales. Aerial and satellite thermal remote sensing require clear sky conditions (clouds are opaque with respect to thermal infrared radiation). Hence, long term satellite characterizations of sUHI are biased towards conditions that maximize urban-rural contrasts in T\textsubscript{surf}. This results in an overestimation of “all-sky” sUHI. At diurnal scales, satellite overpass cycles rarely coincide with sUHI maximums and are not standard across cities or platforms. Thus, analysis of sUHI patterns and magnitudes across cities and instrument platforms is difficult. At smaller time scales still, time-continuous analysis of urban T\textsubscript{surf} shows significant microscale (second to minute) fluctuations in temperature \citep{Christen2012}. Most thermal remote sensors provide instantaneous T\textsubscript{surf} (rather than temporally averaged) and are potentially contaminated by high-frequency microscale fluctuations in urban T\textsubscript{surf}. This is particularly salient in urban environments, where a large variety of fabric materials can produce significant directional contrasts in thermal admittance - and thus spatial variations in the magnitude of microscale fluctuations depending on the facet material types viewed by the sensor. The effect of this phenomenon on thermal remote sensing has not been extensively studied, however, the magnitude of microscale fluctuations in T\textsubscript{surf} is significant relative to a typical sUHI signal and thus constitutes a potentially large source of bias.

@@

First, surface-to-sensor path length and view factor geometries are calculated with the surface-sensor-sun urban model (SUM) \citep{Soux2004}. Using a digital building model, SUM calculates path lengths from the surface to the sensor and angular view factors by projecting the pyrgeometer FOV onto a simplified (orthogonal) representation of the surface, determining which points in the DBM are 'seen' by the sensor, and calculating the distance from each 'seen' point to the sensor. Path lengths are binned at 5\si{\degree} intervals and averaged over the azimuth angle. Finally, view factors are calculated for each 5\si{\degree} angular as each bin's proportion of the total view factor and normalized to unity.

Second, version 4.1 of the MODTRAN radiative transfer code \citep{Berk1987} is used to model spectral at-sensor radiances for each angular path length. Runs are initialized using profiles of T\textsubscript{air} and water vapor content collected concurrently at each time step. Spectral radiances are modeled over a bandpass of 1 to 2500 \si{cm^{-1}} at an emissivity of 0.95 and convolved by an extended dome transmittance function to ensure that modeled radiances replicate the actual remote sensed signal. Consultation with radiometer manufacturers made clear the need to extend the modeled bandpass beyond a typical longwave bandpass (approximately 250 - 2500 \si{cm^{-1}}), as silicone domed pyrgeometers are transmissive of radiation in much shorter wavenumber (longer wavelengths) than 250 \si{cm^{-1}} (40 \si{\micro\meter}).

@@

T\textsubscript{hem, r} are retrieved from radiation and meteorological data collected from December 2001 through July 2002 as a part of the Basel Urban Boundary Layer Experiment (BUBBLE) campaign conducted in Basel, Switzerland \citep{Rotach2005}. Urban data were observed from a tower located near city center at the "Basel Sperrstrasse" site. Rural reference data were observed at the "Lange Erlen" site approximately 6 \si{\kilo \meter} to the north east in the outskirts of the city of Basel. Urban and rural site morphologies and observed variables are described in Table \ref{morphbspr}.

\begin{table}[H]
	\centering
	\caption{A description of morphological and measured variables for urban and rural sites. Modified from \citet{Rotach2005} to include only relevant parameters.}
	\label{morphbspr}
	\begin{tabular*}{\textwidth}{p{3.75cm} p{2.25cm}p{3.5cm}p{2.75cm}p{2.75cm}}
		\toprule 
		Site & Location & Morphological & Meteorological & Radiation \\ 
		& Height & Characteristics\footnote{} & Variables\footnote{} & [No of levels]\footnote{} \\ 	\midrule

		Basel Sperrstrasse \newline \textit{Urban} \textit{street canyon} \newline LCZ: 2 & 47.57\si{\degree} N \newline 7.60\si{\degree} E \newline 255 \si{\meter} a.s.l. & $z_H$ = 14.6  \si{\meter} \newline $\sigma_H $ = 6.9 \si{\meter} \newline H/W = 0.54 \newline $\lambda_C $ = 0.37\newline $\lambda_S$ = 1.0 \newline $\alpha$ = 11.0\% & T\textsubscript{air} [7] \newline H [7] \newline WV [12] \newline WD [1] \newline P [1] & L\textsubscript{up} [3] \newline L\textsubscript{down} [5] \newline S\textsubscript{up} [2] \newline S\textsubscript{down} [3] \\ 
		& & & & \\
		Lange Erlen \newline \textit{Rural parkland} \newline LCZ: B & 47.59\si{\degree} N \newline 7.65\si{\degree} E \newline 240 \si{\meter} a.s.l. & $\alpha$ = 21.4\%  & T\textsubscript{air} [4] \newline H [4] \newline WV [3] \newline WD [1] &  L\textsubscript{up} [1] \newline L\textsubscript{down} [1] \newline S\textsubscript{up} [1] \newline S\textsubscript{down} [1]  \\ 
		\bottomrule
	\end{tabular*} 
		\raggedright
		\textsuperscript{2} Morphological parameters for the Sperrstrasse site were calculated for a 250 \si{\meter} circular area surrounding the study sites using the method described in \citet{Grimmond1999}. $z_H$: average building height, $\sigma_H $: standard deviation of building height, $\lambda_P $: plan aspect ratio, $\lambda_C $: complete aspect ratio, H/W: local canyon height to width ratio, $\alpha$: surface albedo. \\
		\textsuperscript{3} H: humidity, WV: wind velocity, WD: wind direction, P: pressure. \\
		\textsuperscript{4} L\textsubscript{up}: upwelling longwave radiation, L\textsubscript{down}: downwelling longwave radiation, S\textsubscript{up}: upwelling shortwave radiation, S\textsubscript{down}: downwelling shortwave radiation.
\end{table}

@@

Figure \ref{heatsuhi} shows seasonal variations in diurnal patterns of mean monthly hemispherical sUHI magnitudes at 30 minute intervals for the eight month study period. sUHI displays significant seasonal variation, particularly by day, and is strongly controlled by day length and solar angle. Nighttime sUHI does not vary significantly across seasons.

Figure \ref{heatcluhi} shows seasonal variations in diurnal patterns of mean monthly clUHI calculated at 30 minute intervals for the eight month study period. clUHI development is strongly controlled by the timing of sunset/sunrise cycles and displays the largest seasonal variance in the hours immediately after sunset. Daytime clUHI does not vary significantly across seasons. 
\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{heatsuhi}
	\caption{A heatmap of mean half hourly hemispherical sUHI for each month calculated at 30-minute intervals over the eight month study period.}
	\label{heatsuhi}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_kdown}
	\caption{sUHI magnitude versus incoming solar radiation binned for clear sky (when the sum of daytime K\textsubscript{down} \textgreater 10000 \si{\watt \per \square \meter}) and cloudy sky days (when the sum of daytime K\textsubscript{down} \textless 10000 \si{\watt \per \square \meter})}
	\label{meteo_kdown}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_hum}
	\caption{sUHI magnitude versus atmospheric water vapor content measured at 2 \si{\meter} at the Sperrstrasse site. Similar patterns are observed when rural water vapor content is substituted.}
	\label{meteo_hum}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_wv}
	\caption{sUHI magnitude versus rural wind velocity measured at approximately 10 \si{\meter} above ground at the Lange Erlen site.}
	\label{meteo_wv}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_atuhi}
	\caption{sUHI magnitude versus clUHI magnitude. clUHI is calculated as the difference between urban and rural T\textsubscript{air} measured at 2 \si{\meter} above ground level.}
	\label{meteo_atuhi}
\end{figure}

@@

sUHI in Basel from complete, nadir, and hemispherical representations of T\textsubscript{surf} displays two diurnal peaks, the first in the late afternoon and a second just after sunset. Clear sky conditions amplify this phenomena. A late afternoon peak in sUHI is likely the result of the urban-rural contrasts in thermal admittance (and by relation, urban-rural soil moisture contrasts). The thermal admittance of many urban surfaces, particularly rooftops, tends to be lower than rural and vegetated environments \citep{Spronken-Smith1998} – particularly in temperate climates where rural soil moisture is largely maintained through summer. Lower urban thermal admittance manifests in a larger diurnal thermal amplitude and thus, greater daytime surface heating and nighttime surface cooling. Nadir and hemispherical views are biased towards low thermal admittance rooftop facets, and thus display large late afternoon peaks in sUHI.

A second peak in sUHI after sunset is the result of differences in relative cooling rates of urban and rural surfaces. In this case, nighttime sUHI is likely the result of canyon radiation trapping from obstructed canyon sky-view-factors. The second peak sUHI is largest following clear sky conditions when the potential for significant differentials in urban-rural nocturnal cooling rates is maximized. A second peak is observed across all representations of urban T\textsubscript{surf} but is most apparent in complete and nadir representations of sUHI, which, for the Sperrstrasse site, sample T\textsubscript{road} to a greater extent than the hemispherical view. A weak sUHI is maintained through the nighttime as urban and rural cooling rates reach parity in the hours after sunset \citep{Oke2017}. Minimum sUHI is observed just after sunrise, where sUHI magnitudes are occasionally negative. At low solar angles, shading by 3-dimensional urban geometry suppresses heating rates for a large portion of urban surface area, reducing overall urban heating rates in early morning. As solar angles increase through the morning and more facets experience direct isolation, sUHI develops, with nadir and hemispherical representations displaying more rapid intensification than complete sUHI from a bias towards low thermal admittance and directly sunlit rooftop surfaces.

@@

Hemispherical sUHI displays significant seasonality, both in terms of its diurnal pattern and overall magnitude. Seasonal variations in sUHI are largest by day, where large seasonal contrasts in cloud cover frequency and solar angle exert strong control on daytime heating rates and sUHI development. Peak sUHI occurs near the summer solstice, when days are longest and isolation on clear sky days is most intense. In winter, daytime sUHI is suppressed from a lower solar input and frequent cloud cover. This weakens wintertime sUHI and shifts its peak to shortly after sunset. Nighttime sUHI is relatively consistent across seasons. Similar seasonality is observed in clUHI, albeit concentrated at different times of day.

In contrast to sUHI, clUHI is greatest just after sunset and is strongly controlled by sunset/sunrise timing. Peak clUHI occurs just after sunset across seasons, differing only in terms of magnitude, with a larger peak clUHI in summer. However, in this study, it should be note that summertime clUHI magnitudes in the hours just after sunset are slightly higher than expected \citep{Runnalls2000} - possibly the result of cold air draining from hilly terrain west of the rural site. As is the case with sUHI seasonality in clUHI results from seasonal contrasts in solar input and cloud cover frequency. Increased wintertime cloud cover reduces radiative losses by the canyon air volume after sunset, suppressing mean wintertime clUHI magnitudes. Daytime clUHI is consistently small across all seasons. Thus, although sUHI and clUHI are controlled by similar climatic controls and have a similar micro meteorological genesis, they do not display the same seasonal patterns. A surface urban heat island is not simply a stronger, more pronounced, canopy layer urban heat island.

@@

sUHI is highly dependent on meteorological conditions, which account for much of the variation in daytime sUHI magnitudes. Conditions that foster the largest microscale contrasts in urban T\textsubscript{surf} also entail the largest sUHI magnitudes. Contra, conditions that suppress microscale contrasts in T\textsubscript{surf} show similarly reduced sUHI magnitudes. As such, incoming shortwave radiation and wind velocity display positive and negative relationships with sUHI magnitudes respectively. Under clear sky conditions, strong solar heating of the surface combined with differences in urban and rural thermal properties fosters the development of large daytime urban-rural thermal contrasts. Thus, sUHI\textsubscript{max} coincides with insolation\textsubscript{max}. Under cloudy conditions, decreased direct solar heating of the surface suppresses the development of urban-rural thermal contrasts, resulting in a reduced sUHI. $\Delta$T\textsubscript{hem, r - air} has a similar positive effect on sUHI, owing to its dependence on solar insolation.

As was found with correction magnitudes in Chapter \ref{paper1}, sUHI magnitudes are not strongly correlated with water vapor content as, in the absence of fog, humidity near the surface does not strongly influence surface heating or cooling rates in and of itself. Atmospheric transmittance of longwave radiation decreases sharply with small increases in water vapor when the atmosphere is very dry - shown in Figure \ref{humtest}. As water vapor content increases from zero, variation in atmospheric transmittance decrease. Therefore, typical variations in urban and rural humidities have little effect on the fraction of outgoing available for heating of the canopy layer air volume through absorption of outgoing longwave radiation and for heating of the surface by atmospheric emission. As a result, water vapor content has little effect on sUHI. Wind velocity displays a strong negative relationship with sUHI. sUHI is greatest under calm winds when near-surface mixing is minimal, fostering strong microscale and urban-rural contrasts in T\textsubscript{surf}. As wind velocity increases, sUHI is suppressed from increased turbulent heat losses. Meteorological controls on sUHI are summarized in Table \ref{meteo_cont}.

\begin{table}[H]
	\centering
	\caption{A summary of the relationship between sUHI, meteorological variables, and clUHI.}
	\label{meteo_cont}
	\begin{tabular}{lc}
		\toprule 
		An increase in $x$, has a/an... & $y$ effect on sUHI magnitude. \\
		\midrule
		$\Delta$T\textsubscript{hem, r - air} &\textit{increasing} \\
		&\\
		T\textsubscript{hem, r} & \textit{increasing} \\
		&\\
		T\textsubscript{air} & \textit{increasing} \\

		&\\
		Incoming solar radiation & \textit{increasing}  \\
		& \\
		& \\
		Water vapor content & \textit{negligible} \\
		&\\
		Wind velocity &\textit{decreasing} \\
		&\\
		clUHI &\textit{increasing} \\
		\bottomrule
	\end{tabular} 
\end{table}

clUHI and sUHI develop at different times of day and, thus, have a complex diurnal relationship. By day, variation in clUHI is minimal and shows a weak positive correlation with sUHI. By night, clUHI shows greater variation and a stronger, more positive relationship with sUHI. The relationship between clUHI and sUHI is most evident on clear sky nights when urban-rural contrasts in air and surface cooling rates are strongest. Under these conditions sUHI and clUHI  show similar development in the hours after sunset.
